The evolution of gene expression in the term placenta of viviparous mammals

### 2016-03-25 manuscript/manuscript.Rnw
fix typos and make it log_2 FC

All of the code used to produce every analysis presented in this paper
is available at
\url{https://github.com/uiuc-cgm/placenta-viviparous.git}. All of the
sequences used are deposited in SRA, and new sequences for this work
can be obtained under the series
\href{http://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE79121}{GSE79121}
(Supplemental \cref{tab:srr_numbers}). The mean FPKM for gene in each
species with ortholog annotations are located in the supplemental data
files \Sexpr{english.join(paste0(gsub(" ","_",
  star.logs[,unique(species)]),"_mean_fpkm.txt"))}.

@@

        display.table <-
            display.table[,`Human Symbol`:=paste0("\\textit{",xtable.sanitize.text(`Human Symbol`),
                                                  "}")]
        print(xtable(display.table,
                     display=c("s","s","s","e","g","f"),
                     align=c("c","c","p{2in}","r","r","r")
                     ),
              booktabs=TRUE,
              floating=FALSE,
              math.style.exponents=TRUE,
              sanitize.text.function=genes.table.sanitize,
              include.rownames=FALSE)

@@

Two genes, \textit{NPPB} and \textit{NPPC}, code for peptide hormones
that are important in the inducement of vasodilation, and may be
important in increasing maternal blood flow to the fetus
\cite{Cameron.ea1996:sitesofgeneexpression}. There is some indication
that these proteins are present in the decidua of mice
\cite{Cameron.ea1996:sitesofgeneexpression}, but we find very little
expression in the fetal portion of the mouse placenta
(\cref{fig:hormone_genes}, Supplemental
\cref{fig:nppc_evolution_tree,fig:nppb_evolution_tree}). While we do
find high expression of at least one of these genes in \dog, \cow, and
\elephant, the expression pattern does not follow the placental
organization (endotheliochorial for \dog and \elphant,
synepitheliochorial for \cow and \sheep), so it is unlikely that
expression of these genes are essential for specific placental
organizations.

>>>>

All of the code used to produce every analysis presented in this paper
is available at
\url{https://github.com/uiuc-cgm/placenta-viviparous.git}. All of the
sequences used are deposited in SRA, and new sequences for this work
can be obtained under the series
\href{http://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE79121}{GSE79121}
(Supplemental \cref{tab:srr_numbers}). The mean FPKM for gene in each
species with ortholog annotations are located in the supplemental data
files \Sexpr{english.join(paste0(gsub(" ","\\_",
  star.logs[,unique(species)]),"_mean_fpkm.txt"))}.

@@

        display.table <-
            display.table[,`Human Symbol`:=paste0("\\textit{",xtable.sanitize.text(`Human Symbol`),
                                                  "}")]
        setnames(display.table,"Fold Change","$\\mathrm{log}_2$ FC")
        print(xtable(display.table,
                     display=c("s","s","s","e","g","f"),
                     align=c("c","c","p{2in}","r","r","r")
                     ),
              booktabs=TRUE,
              floating=FALSE,
              math.style.exponents=TRUE,
              sanitize.text.function=genes.table.sanitize,
              include.rownames=FALSE)

@@

Two genes, \textit{NPPB} and \textit{NPPC}, code for peptide hormones
that are important in the inducement of vasodilation, and may be
important in increasing maternal blood flow to the fetus
\cite{Cameron.ea1996:sitesofgeneexpression}. There is some indication
that these proteins are present in the decidua of mice
\cite{Cameron.ea1996:sitesofgeneexpression}, but we find very little
expression in the fetal portion of the mouse placenta
(\cref{fig:hormone_genes}, Supplemental
\cref{fig:nppc_evolution_tree,fig:nppb_evolution_tree}). While we do
find high expression of at least one of these genes in \dog, \cow, and
\elephant, the expression pattern does not follow the placental
organization (endotheliochorial for \dog and \elephant,
synepitheliochorial for \cow and \sheep), so it is unlikely that
expression of these genes are essential for specific placental
organizations.
